---
title: "Class 12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup for Docking

We will first prepare our HIV-Pr system for drug docking by making a protein only PDB formate file (remove water, existing ligand, etc)

```{r}
library("bio3d")
```

```{r}
file <- get.pdb("1hsg")
```

```{r}
pdb <- read.pdb(file)
pdb
```


Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure?

```{r}
prot <- atom.select(pdb, "protein", value=TRUE)
write.pdb(prot, file="1hsg_protein.pdb")
prot
```

```{r}
lig <- atom.select(pdb, "ligand", value=TRUE)
write.pdb(lig, file="1hsg_ligand.pdb")
lig
```

We will load these into ADT to add hydrogens and make PDBQT files for each

## Process our docking results

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```


## Normal Mode analysis for flexibility prediction

```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```


















