---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database

The [PDB](http://www.rcsb.org/) is the main repository for biomolecular structure data.

Here we examine the contents of the PDB:

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
db <- read.csv("Data Export Summary.csv", row.names=1)
head(db)
```

How many are X-ray?

```{r}
(db$Total[1]/sum(db$Total)) * 100
```

How many are protein structures?

```{r}
(sum(db$Proteins)/sum(db$Total)) * 100
```

We could also try the datapasta package and copy from website and "Addins" > Paste as data.frame ...

```{r}
library(datapasta)

data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "Other", "NMR", "Multi Method",
                           "Electron Microscopy", "Total"),
              Proteins = c(126880, 256, 11062, 129, 2277, 140604),
         Nucleic.Acids = c(2012, 4, 1279, 5, 31, 3331),
    Protein.NA.Complex = c(6547, 6, 259, 2, 800, 7614),
                 Other = c(8, 13, 8, 1, 0, 30),
                 Total = c(135447, 279, 12608, 137, 3108, 151579)
)


```

Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 1157 as of 2019-05-07 See:
http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=5C35D458


## Section 3 Using Bio3D

```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg")
```

```{r}
pdb
```

 Q6. How many amino acid residues are there in this pdb object and what are the two non- protein residues?
 
 There are 198 amino acids. The two non-protein residues are HOH (127) and MK1 (1)
 
```{r}
attributes(pdb)
```
 
```{r}
head(pdb$atom)
```

```{r}
pdb$atom[1:2, c("eleno", "elety", "x", "y", "z")]
```


```{r}
pdb$atom$elety[1:2]
```


```{r}
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab=
            "B-factor")
```


 Q7. What type of R object is   ? HINT: You can always use the str() function to get a useful summery of any R object.
 
```{r}
pdb$xyz
```
 
```{r}
dim(pdb$xyz)
```

```{r}
pdb$xyz[ 1, atom2xyz(1:2) ]
```

### Atom Selection

```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

```{r}
prot.pdb <- atom.select(pdb, "protein", value=TRUE)
write.pdb(prot.pdb, file="1hsg_protein.pdb")
```

```{r}
lig.pdb <- atom.select(pdb, "ligand", value=TRUE)
write.pdb(prot.pdb, file="1hsg_ligand.pdb")
```

Section 5: Example Application on Adenylate Kinase (Adk)

```{r}
aa <- get.seq("1ake_A")
```

```{r}
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```


```{r}
head(hits$pdb.id)
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
pdbs <- pdbaln(files)
ids <- basename.pdb(pdbs$id)
plot(pdbs, labels=ids)
```















