---
title: "Class 6 HW"
author: "Gabrielle Blizard"
date: "4/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

## Q1: What type of object is returned from the read.pdb() function?

### The read.pdb function is used to access the on-line PDB file for proteins and makes a "PDB" object.

## Q2: What does the trim.pdb() function do?

### The trim.pdb function is used to create a smaller PDB object with only a subset of the atoms.

## Q3: What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?

### We could turn off the marginal black and grey rectangles by removing the sse parameter from the plotb3 code, which is preset to = NULL. The black and grey rectangles represent the secondary structure elements of the protein.

## Q4: What would be a better plot to compare across the different proteins?

### A cluster dendogram.

## Q5: Which proteins are more similar to each other in their B-factor trends. How could you quantify this?

### Protein s1.b and protein s3.b are more similar to each other in their B-factor trends. This can be quantified using a cluster dendogram.
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

##Q6: How would you generalize the original code above to work with any set of input protein structures?

### What functions are we repeating in the original code?

#### read.pdb() allows us to access an online protein data bank
*s1 <- read.pdb("4AKE")*

#### trim.pdb allows us to produce a smaller subset of the PDB
*s1.chainA <- trim.pdb(s1, chain="A", elety="CA")*

#### this command allows us to select for specific data in our data set
*s1.b <- s1.chainA$atom$b*

#### plotb3 allows us to plot our protein sequence
*plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")*

### How can we make this code a function?

```{r}
protein_plot <- function(pdbid="", chain="", elety="") {
  
  # read protein data bank
  protein <- read.pdb(pdbid)
  
  # create a smaller PDB object containing specified chain and atoms
  protein_chain <- trim.pdb(protein, chain = chain, elety = elety)
  
  # select data from dataset
  protein_chain_B <- protein_chain$atom$b
  
  # plot protein B factor
  plotb3(protein_chain_B, sse = protein_chain, typ="l", ylab="Bfactor")
}
```

### What are our inputs?
#### There are three inputs to the function:
  *pdbid,*
  *chain,*
  *elety*
  
### What is our output?
#### The output of this function is a *plot of the PDB file B factor vs Residue*

### Try it out for protein "4AKE"

```{r}
protein_plot("4AKE", "A", "CA")
```

### Try it out for protein "1AKE"

```{r}
protein_plot("1AKE", "A", "CA")
```

### Try it out for protein "1E4Y"

```{r}
protein_plot("1E4Y", "A", "CA")
```






