---
title: 'Class 9:Unsupervised Learning Analysis of Human Breast Cancer Cells'
author: "Gabrielle Blizard"
date: "4/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparing the data

## Open the data set
```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
```

## View data set
```{r}
head(wisc.df)
```

## Convert features of the data to a matrix
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
```

## Assign row names
```{r}
row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```

## Store the diagnisis data as a vector of 1 and 0 values with 1 being cancer
```{r}
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```

Q1: How many patients are in our data set?
```{r}
nrow(wisc.df)
```

Q2: How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.data)
```

```{r}
length( grep("_mean", colnames(wisc.data)) )

# colnames(wisc.data)
```

Q3: How many observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```

# Performing PCA

Check if the data needs to be scaled
```{r}
round( colMeans(wisc.data), 1 )

```

```{r}
round( apply(wisc.data, 2, sd), 1 )
```

Execute PCA
```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1, xlab="PC1", ylab="PC2")
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
```{r}
summary(wisc.pr)
```

# Interpreting PCA results

Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

```{r}
biplot(wisc.pr)
```

Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col=diagnosis+1, xlab="PC1", ylab="PC3")
```

# Variance explained

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
pve <- pr.var/30
```

Plot variance explained for each principal component
```{r}
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = "o")
```

Alternative scree plot of the same data, note data driven y-axis
```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

Plot cumulative proportion of variance explained
```{r}
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

Setup side-by-side plots with par(mfcol=c(1,2))
```{r}
par(mfcol=c(1,2))

plot( pve , xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")

plot( cumsum(pve) , xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

# Hierarchial Clustering 

Scale wisc.data
```{r}
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust(data.dist, method = "complete")
```

Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```

# Clustering on PCA results

```{r}
wisc.pca.hclust <- hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2" )
```

```{r}
plot(wisc.pca.hclust)
```

```{r}
grps <- cutree(wisc.pca.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis+1)
```

# Prediction

Read new data
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```


plot new cell data onto our PCA
```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16)
```

Q17. Which of these new patients should we prioritize for follow up based on your results?

We should prioritize the patient that falls in the black data points





