---
title: "Class 8: Machine Learning"
author: "Gabrielle Blizard"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## K-means clustering

Let's start with an example of the **kmeans()** function

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

```{r}
# Use the kmeans() function setting k to 2 and nstart=20
km <- kmeans(x, 2, 20)
```

```{r}
#Inspect/print the results
print(km)
```

> Q. How many points are in each cluster?
There are 30 points in each cluster

> Q. What ‘component’ of your result object details

```{r}
#cluster size?
km$size
```
      
```{r}
#cluster assignment/membership?
km$cluster
```

```{r}
#cluster center
km$center
```


```{r}
#Plot x colored by the kmeans cluster assignment and add cluster centers as blue points
plot(x, col=km$cluster)
points(km$centers, pch=18, col="blue", cex=3)
```


## Hierarchial Clustering example

We must gie the **hclust()** function a distance matrix not the raw data as input

```{r}
# Distance matrix calculation
d <- dist(x)

# Clustering
hc <- hclust(d)
plot(hc)
```

```{r}
cutree(hc, h=6)
```

```{r}
cutree(hc, k=2)
```

 
```{r}
 # Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

```{r}
# Step 2. Plot the data without clustering
plot(x)
```

```{r}
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```


>Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters

>Q. How does this compare to your known 'col' groups?

```{r}
hc <- hclust(dist(x))
plot(hc)
abline(h=2, col="red")
abline(h=2.8, col="blue")
```

```{r}
gp2 <- cutree(hc, k=2)
gp3 <- cutree(hc, k=3)

```

```{r}
gp2
```

```{r}
gp3
```

```{r}
plot(x, col=gp3)
```

```{r}
table(gp2)
```

```{r}
table(gp3)
```

```{r}
table(gp2, gp3)
```



## PCA: Principal Component Analysis

We will use the **prcomp()** function for PCA

```{r}
## You can also download this file from the class website!
     mydata <- read.csv("https://tinyurl.com/expression-CSV",
                        row.names=1)
     head(mydata, 10)
```

100 genes in the dataset
```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

```{r}
colnames(mydata)
```

```{r}
pca <- prcomp(t(mydata), scale = TRUE)
```

```{r}
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```

```{r}
## Percent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
```

```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE) ## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2])
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```


```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
barplot(pca.var.per, main="Scree Plot",
        xlab="Principal Component", ylab="Percent Variation")
```


```{r}
  ## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
   
```

```{r}
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
          xlab=paste0("PC1 (", pca.var.per[1], "%)"),
          ylab=paste0("PC2 (", pca.var.per[2], "%)"))
     ## Click to identify which sample is which
     identify(pca$x[,1], pca$x[,2], labels=colnames(mydata))
```

## PCA of UK food data

```{r}
x <- read.csv("UK_foods.csv")
head(x)
```

Q1

```{r}
nrow(x)
```

```{r}
ncol(x)
```

Q2

```{r}
x <- read.csv("UK_foods.csv", row.names=1)
head(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Q3: changing beside = FALSE

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

Q5: if a point lies on the diagonal it means the two variables being compared are exactly the same.

```{r}
pairs(x, col=rainbow(10), pch=16)
```

Q6

```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(x) )
summary(pca)
```

Q7: PC1 vs PC2

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))

```

Q8

```{r}
colnames(x)
```

```{r}
mycols <- c("red", "orange", "blue", "green")
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=mycols)

```


## Digging deeper (variable loading)

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )

```


















