---
title: "Class 14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import countData and colData

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors=FALSE, row.names=1)
head(counts)
```

```{r}
metadata <- read.csv("airway_metadata.csv", stringsAsFactors=FALSE)
head(metadata)

```

There are `r nrow(counts)` in this dataset

```{r}
nrow(counts)
```

Lets check the corespondence of our countData cols and our mataData rows. This will be required when we use DESeq later!
```{r}
colnames(counts)
```

```{r}
metadata$id
```

```{r}
all( colnames(counts) == metadata$id )
```

## Section 3: Walk through of gene expression analysis

Our metadata file tells us which columns of the countData are drug treated and control (i.e. non drug treated)

```{r}
control.inds <- metadata$dex == "control"
control <- metadata[control.inds,]
control 
```

```{r}
head( counts[, control$id] )
```

```{r}
control.mean <- rowSums(counts[, control$id]) / length(control$id)
```

Do the same for the drug treated experiments...

```{r}
treated.inds <- metadata$dex == "treated"
treated <- metadata[treated.inds,]
treated
```

```{r}
treated.mean <- rowSums(counts[, treated$id]) / length(treated$id)
```

Let's store the control.mean and treated.mean together for ease of use
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

What is the sum counts across all genes in control and treated

```{r}
colSums(meancounts)
```

```{r}
plot(meancounts)
```

Plot on a log-log scae so we can see our data
```{r}
plot(meancounts, log="xy")
```

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the head() or the View() function for example.

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

```{r}
to.rm <- unique(which( meancounts[,1:2] == 0, arr.ind=TRUE)[,1] )
```

```{r}
mycounts <- meancounts[-to.rm,]
nrow(mycounts)
```

How many genes are up in the drug treated cells?
```{r}
up.inds <- mycounts$log2fc > 2
sum(up.inds)
```

and how many are down?

```{r}
down.inds <- mycounts$log2fc < -2
sum(down.inds)
```

```{r}
head( mycounts[up.inds,])
```

##Section 4: Adding annotation data

We can add annotation from a supplied CSV file, such as those available from ENSEMBLE or UCSC. The annotables_grch38.csv annotation table links the unambiguous Ensembl gene ID to other useful annotation like the gene symbol, full gene name, location, Entrez gene ID, etc.

```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

use the **merge()** function to aff the annotation data from the `anno` object to our RNA-Seq results in `my counts`
```{r}
# use the merge function
mycounts.anno <- merge(mycounts, anno, by.x="row.names", by.y="ensgene")
mycounts.anno
```


Install the bioconductor annotation packages
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

We will use the **mapIds()** function from bioconductor to add annotation data to our results

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db, 
                   keys=row.names(mycounts), # the gene names in our dataset
                   keytype="ENSEMBL", # format of our gene names
                   column="SYMBOL") # what new data do we want to add
```

```{r}
head(mycounts)
```

```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db, 
                   keys=row.names(mycounts), # the gene names in our dataset
                   keytype="ENSEMBL", # format of our gene names
                   column="ENTREZID") # what new data do we want to add
```

```{r}
head(mycounts)
```

##Section 5 

```{r}
library("DESeq2")
```

Setup the object needed for DESeq analysis

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors=FALSE)
```


```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
head(res)
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r}
res01 <- results(dds, alpha=0.01)
head(res01)
```


Let's sort or order our results by the adjusted p-value

```{r}
x <- c(5,4,1,2)
x[ order(x) ]
```

```{r}
ord.inds <- order( res01$padj )
write.csv( res01[ord.inds, ] , file="signif01_results.csv")
```

## Volacano plot

```{r}
plot(res01$log2FoldChange, -log(res01$padj))
```


```{r}
mycols <-  rep("gray", nrow(res01))
mycols[ res01$padj < 0.01 ] <-  "black"
#mycols[ abs(res01$log2FoldChange) > 2] <- "Blue"

mycols[ (res01$padj < 0.01) & (abs(res01$log2FoldChange) > 2)] <- "red"

plot(res01$log2FoldChange, -log(res01$padj), 
     xlab="Fold Change (Drug/Control",
     ylab="-log(P-value)",
     col=mycols)
```










